Hysteresis in human computation:\\ how one task affects another

### 2014-11-19 docs/publications/2014.09_SCIENCE/man.tex
spell check

People still outperform sophisticated artificial 
intelligince systems in many tasks requiring general knowledge, 
qualitative judgment, and visuo-spatial reasoning. Recently, microtask 
platforms enable the programmer to import human judgment directly into 
algorithms: programs can post tasks, to be completed by users of such 
platforms in near real-time.
Some computer scientists consider microtask platforms to be a fundamentally
new computing architecture, but psychological effects can introduce systematic
bias. 
Here we investigate the effects that earlier tasks have on the results of 
subsequent ones, which we call \textit{inter-task effects}.
In a canonical image labeling task, we found that inter-task effects are
very strong---stronger than overtly framing tasks in terms of 
particular research goals.  However, in some cases, inter-task effects can
be leveraged to elicit more nuanced responses.
To study inter-task effects we introduce a framework that characterizes HPU 
bias in an application-independant way, based on its potential to alter an 
algorithm's path of execution.  Our findings suggest that careful consideration
should be given to inter-task effects when designing crowdsourcing studies.

%microtask worker performs a series of tasks.  We find that early tasks can
%have a strong effect on later tasks, and that these effects can be stronger
%than those brought about by framing.
%
%course of human computation algorithms.
%application-independant quantity that measures the
%severity of priming.  We apply this framework to measure 
%
%in many
%tasks, especially ones requiring general knowledge about the world, qualitative
%judgment, and visuo-spatial reasoning.  Recently, microtask web platforms have 
%emerged. One can upload a batch of tasks to be completed by users who are paid 
%for their efforts.
%Though simple on the surface, these platforms dramatically enlarge the
%programmers toolkit: in effect they allow the programmer to import human 
%judgment into algorithms.  Some computer scientists consider
%such platforms to be a new computing architecture.  But people are very 
%different from electronic processors.  For example, context affects how a 
%person will respond to a prompt, and this complicates the formal treatment of 
%human computation.  To approach this issue, we introduce a framework 
%that quantifies the severity of priming in an application-independant way.
%We apply this technique to reveal an unforseen source of bias: the bias 
%induced by one task on another.  Surprisingly we find that 
%\textit{inter-task effects} introduce more bias than \textit{framing}, which
%has received more attention.
%
%more well-known 
%\textit{framing}, a more well-studied form of priming.
%
%Microtask platforms offer a fast, flexble, and inexpensive source of labour for
%clerical tasks \cite{Finnerty2013, chandler2013breaking, Berinsky2012351} and 
%recently have been adopted by researchers to augment datasets and solicit 
%participants \cite{paolacci2010running, Berinsky2012351, chandler2013breaking}.
%Computer scientists have described such platforms as a new computing 
%architecture, and seek to characterize “human co-processing units” (HPUs) in 
%analogy to CPUs and GPUs \cite{5543192}. However, it is known that people are 
%susceptible to 
%priming effects\cite{No2007,Swaab200299,Gopher2000308,sohn2001task,Ghuman17062008,BJOP1796,BJOP1826,Gass1999549}, which stands to complicate the HPU 
%perspective.  Here, we introduce a framework to recast this psychological 
%phenomenon as a kind of computational hysteresis. 
%We apply this framework to study \textit{inter-task effects} that arise when 
%microtask 
%workers perform a sequence of tasks. Surprisingly, earlier tasks can have a 
%very strong influence on workers during later tasks---stronger than overtly
%framing the tasks as serving different research goals.   
%This result has major implications for the design of crowdsourcing initiatives.
%The framework we introduce provides a general measure of priming severity, 
%and, in combination with machine learning tools, enables measurement without
%knowing ahead of time how priming will affect output.
\end{sciabstract}

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of small tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental 
participants in a research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results \cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}

These platforms make it possible to seemlessly integrate human and 
machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Attempts have been made to establish an HPU instruction set.  For example,
the X library provides function calls such as 
\texttt{X.vote()}, and \texttt{X.map()}
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Of course, there are obvious differences between HPUs and their silicone
counterparts.  When facing the same task, people do not generally
respond in the same way.  Common sense would dictate that a certain amount
of variability in responses made by people is inescapable.  But an important 
part of the variation is bound to be systematic, due to psychological or 
demographic factors.

Here, we highlight an important source of bias with
serious implications for any crowdsourcing initiative. 
It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{inter-task} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs.

There has been considerable investigation into the factors that affect the 
output from microtask work.  These include such factors as the 
level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
inter-task effects on microtask platforms.

We make two contributions towards understanding inter-task effects,
and systematic bias in HPUs in general.  First, we present a framework for 
quantifying bias in HPU outputs in terms of their potential impact 
algorithms.  This yields an approach for measuring the severity of bias 
that removes the need to understand the underlying psychological or 
demographic factors at play.  Second, we apply our framework to measure the
severity of inter-task effects on a canonical microtask: image labeling.

Using the  Amazon Mechanical Turk (Mturk) platform, we solicited workers to 
perform a series of image-labeling tasks.  We chose image labeling because it 
is one of the most common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
of applications in computer vision research \cite{5543192}.
%In these tasks, workers labeled images featuring iconic cultural scenes,
%food, table-settings, and various utensils and cookware. 
We divided the tasks into an \textit{initial} and \textit{test} set, 
but, crucially, there was no distinction between these sets from the
perspective of the worker.  We varied the initial tasks, while 
keeping the test set the same, to analyze the effect of initial tasks on 
the labels provided during the test tasks.

As a point of comparison, we included treatments in which we induce 
\textit{framing}.
Before the workers labeled images, we indicated that the work 
was funded by a research lab with a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. 
Surprisingly, inter-task effects were stronger than framing, in some cases 
about twice as strong. 
Our results demonstrate that initial tasks can change workers' focus in later 
tasks.  We also found that inter-task effects altered the level of specificity 
of worker's labels.  Our findings suggest that careful consideration should be 
given to the bundling of tasks when designing a study using a microtask 
platform.

\paragraph*{Computational hysteresis.}
Before describing our experiments, we dedicate some space to discussing
how we characterize and measure priming in HPU outputs.  Since HPU outputs are
stochastic, determining the effect of priming on an HPU population
amounts to determining the resulting change to the HPU population's 
distribution of outputs.

To refine this, we distinguish between methods that determine \textit{whether}
two random sources have the same distribution 
(e.g. Pearson's chi squared test), from those that determine
\textit{the extent} of that difference (e.g. one of the many measures of 
statistical divergence).  We are interested in measuring the extent of
difference because our practical concern is the potential for such
differences to alter the course of an algorithm.

To measure the effect of some priming treatment, we must of course draw a 
comparison to some standard population.  In many cases, there may be an 
obvious comparison between a primed and unprimed population.  However, we 
refrain from proposing a canonical neutrally-primed HPU population; what is 
considered ``unprimed'' in one scenario may be ``primed'' in another. 
Instead, we speak of two HPU populations being \textit{differently primed}.

We present three alternative, equivalent definitions for the 
\textit{difference in priming} of two HPU populations, $\mathcal{P}$, and 
$\mathcal{Q}$, which we denote $\theta(P,Q)$.  To set up the definitions,
we will assume that $\mathcal{P}$ and $\mathcal{Q}$ 
were sampled randomly from the same initial population $\mathcal{R}$, and 
then were exposed to (possibly) different priming treatments. Then, the HPUs 
in $\mathcal{P}$ and $\mathcal{Q}$ perform some task, leading to popualtions 
of work-products, $P$ and $Q$.

The first definition provides the motivation for our choice of divergence
measure, because it directly measures the proportion of the time that a
binary decision will be changed due to priming effects:

\begin{mydef}
	\label{def:bias}
	{\upshape Maximum Bias:}
	Let $\mathcal{D}$ be a binary decision function, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta_\mathcal{D}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ and $Q$:
	$$
	\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	\theta_\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $\mathcal{P}$ and $\mathcal{Q}$ to be the worst-case bias, i.e. 
	the bias of the most susceptible decision:
	$$
	\theta(P,Q) = 
		\sup_\mathcal{D} \theta_\mathcal{D}(P,Q)
	$$
\end{mydef}
If $\mathcal{D}$ is the condition of an \texttt{if}-statement, then the 
definition above is the change in the fraction of the time that a given 
path of execution is followed due to HPU priming.

While this is precisely what we want to measure, it is not at all clear how
to go about doing it.  The second formulation provides a practical connection
by rephrasing priming in terms of its distinguishability:

\begin{mydef}
	{\upshape Distinguishability:}
	Let $\mathcal{D}$ be a distinguisher algorithm, which takes work-product 
	$r \in P \cup Q$ as input, and outputs a \emph{guess} 
	$\mathcal{D}(r) \in \{0, 1\}$. Now define a validation test 
	$V(\mathcal{D}, R_0, R_1)$ as follows. A bit is drawn uniformly at random, 
	$b \in_R \{0, 1\}$, and then a work-product $r$ is chosen uniformly at 
	random from one of the populations, decided by the bit: $r \in_R R_b$. 
	Input $r$ into $\mathcal{D}$, and if $\mathcal{D}(r) = b$, we say 
	$\mathcal{D}$ \emph{guessed correctly}, and the value of $V$ is 1, 
	otherwise $V = 0$.

	A distinguisher that guesses randomly will still be correct half of the 
	time. Therefore, denote the performance of $\mathcal{D}$ by 
	$\eta_\mathcal{D}(P, Q)$, and let
	$$
		\eta_\mathcal{D}(P,Q) = 2\cdot \mathrm{E}\{V(\mathcal{D},P,Q)\} - 1.
	$$
	Then define the difference in priming between $\mathcal{P}$ and 
	$\mathcal{Q}$ to be the performance of the best possible distinguisher:
	$$ 
		\theta(P,Q) = \sup_\mathcal{D} \eta_\mathcal{D}(P,Q)
	$$
	\label{def:dist}
\end{mydef}

\textbf{Definition 2} relates the difference in priming to the 
\textit{optimal Bayes classifier} for determining whether a given HPU output
came from $P$ or $Q$.  This means that we can obtain a lower bound on
$\theta(P,Q)$ based on the performance of an actual classifier, which provides
recourse when purely statistical methods fail.

The statistical divergence measure underlying the previous definitions is the
divergence measure known as the ``total variational distance'' or 
$l_1$-distance:
\begin{mydef}
	\label{def:l1}
	{\upshape Total variational distance $\theta_{l_1}$:}
	Define the difference in priming between $\mathcal{P}$ and $\mathcal{Q}$ 
	to be equal to the total variational distance between the work-products 
	$P$ and $Q$:
	$$
	\theta_{l_1}
	= \sup_{A \subset \mathcal{X}} \left| \sum_{x \in A} f_P(x) - f_Q(x)\right|
	= \frac{1}{2} \sum_{x \in \mathcal{X}} \left| f_P(x) - f_Q(x) \right|,
	$$
	where $f_P(x)$ denotes the probability of observing x when sampling from P.
\end{mydef}
The equivalent formulation for $\theta(P,Q)$ in 
\textbf{definition \ref{def:l1}} provides a grounding in a fundamental
measure of statistical divergence, and a connection to information theory.  
The equivalence of the three definitions
follows from the known fact that no algorithm can have output that differs
on inputs $P$ and $Q$ more than $l_1 \times 100\%$ of the time.  We provide
proofs in the supplementary material.

The difficulty of measuring $\theta(P,Q)$ depends on what the universe of
possible outputs are, as well as how $P$ and $Q$ are actually distributed,
both of which are mainly determined by the task that the HPUs are asked to 
perform.  In the case where workers label an image with a free-text input, 
the number of different possible labels can be in the thousands or more.  If
the support, $\mathcal{X} = P \cup Q$, is large compared to the number of 
HPU outputs sampled, so that many labels appear only once (which in any case
is expected from Zipv's law), then estimating $\theta$ directly from the
observed frequencies of labels is difficult.

One approach to estimating the $l_1$-distance from samples, which appears
reasonable on the surface, is to use the observed frequencies of each output
$x$ from $P$ and $Q$ to estimate $f_P(x)$ and $f_Q(x)$, and calculate 
$\theta(P,Q)$ directly from \textbf{definition 3}.  This is the so-called naive
approach to estimating $l_1$-distances, and is aptly named because it tends
to produce very inflated values.  More recent techniques have for estimating
$l_1$ have been suggested by Valiant \cite{val-thesis}, and Batu 
\cite{batu2000testing,batu2013testing}.  We reserve the discission of these
techniques to the supplementary material; they do not turn out to be useful
because they do not provide confidence intervals.  Thus our only recourse is
to test the performance of classifiers at distinquishing $p \in P$ from 
$q \in Q$, which allows us to estimate a lower bound on $\theta(P,Q)$.  This 
turns out to be sufficient, and has the advantage that, so long as they are
learnable, it is not necessary to understand exactly how $P$ and $Q$ differ.

\paragraph{Experimental setup.}
We performed two experiments, soliciting respectively 900 and 2300 MTurk 
workers to label images depicting iconic cultural scenes, meals, ingredients, 
table-settings, and various utensils and cookware. In these experiments,
workers were randomly assigned to treatments, and depending on this 
assignment, were exposed to different framing primes, given a different 
set of initial images, and shown a different permutation of the test images. 
For both experiments, workers were first shown brief instructions.  
The framing prime was then shown, unless the given 
treatment involved no framing.  Workers then performed a series of 
image-labeling tasks and had to provide five labels per image. 
For the purpose of analysis, we denote the first five images as the 
\textit{initial set}, and the last five as the \textit{test set}, but no
distingction was made from
the perspective of the worker.
Table 1 sumarizes the various experimental treatments for both 
experiments.

\begin{table}[t]
\centering
	\begin{tabular}{ l  l  l }
		\hline                       
		Treatment & Framing prime & Initial image set	\\ 
		\hline                       
		\textit{Experiment 1}\\
		$\textsc{ambg}$ & None & Ambiguous\\
		$\textsc{img:cult}$ & None & Cultural\\
		$\textsc{img:ingr}$ & None & Ingredients\\
		$\textsc{frm:cult}$ & Cultural & Ambiguous\\
		$\textsc{frm:ingr}$ & Ingredients & Ambiguous\\
		\hline  
		\textit{Experiment 2}\\
		$\textsc{img:food}$ & None & Food\\
		$\textsc{img:obj}$ & None & Objects\\
		$\textsc{frm:food}$ & Food & None\\
		$\textsc{frm:obj}$ & Objects & None\\
		\hline  
	\end{tabular}

	\caption{ \footnotesize{ 
		Workers were uniformly randomly assigned to one of the 
		treatments listed above. 
		The full funder names used were 
		Cultural: ``The Global Foundation for Cultural Recognition'' and 
		Ingredients: ``The National Foundation for Nutritional Awareness''.  
		The ambiguous, cultural, and ingredients initial image sets are shown 
		in \textbf{Figs. 2}, \textbf{3}, and \textbf{4}.
	}}
	\label{table:1}
\end{table}

\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/tasks.pdf}
	\caption{Examples of slides used for the image labeling tasks. A) instructions shown to all workers; B) a framing slide; C) a slide from the initial set for treatment IMG:FOOD; D) a slide from the initial set for treatment IMG:OBJ; E) a slide from the test set shown in all treatments. The full set of slides used for all treatments is presented in the supplementary material.}
	\label{fig:task}
\end{figure}

>>>>

Microtask platforms combine the efforts of large numbers 
of people to perform tasks that are difficult to automate with computers 
alone.  These platforms support near-real-time task completion, mimicking a
compute server, and shift the role of the human from consumer to 
producer of computing resources.  To realize its full
potential, we must understand how human computation differs from machine 
computation, and address the biases inherent in 
human cognition.  The effects of various task-design variables on worker
performance have been investigated, but there has been little 
investigation into how consecutive tasks influence one another. Here we show 
that the responses in later tasks are strongly influenced by the content of 
earlier ones, an effect that is as strong or stronger than the effect of 
framing.  Our findings are not solely cautionary; we find
that these \textit{intertask effects} can be harnessed to improve the quality 
of responses, and might thereby help to achieve expert-level judgments in 
citizen science and other crowdsourcing initiatives. 
As an ancillary contribution, we demonstrate the use of machine learning 
algorithms measure exposure-induced human bias, in terms of its ability to 
influence human computations. 
\end{sciabstract}

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental participants in a 
research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
This trend has seeded considerable interest in characterizing the parameters
of task design, and validating the microtask platform methodology (\#).

Computer scientists view microtask platforms in the context of general
computation.  In this regard, the platforms enable a tight integration of 
human and machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag} (\#).  Libraries 
are being developed to give the programmer more seamless access to HPU 
resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
As with other architectures, the HPU architecture makes entirely new classes 
of problems tractable for the programmer (\#).

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy.  Part of this noise can be explained
in terms of worker demography (\#).  But systematic cognitive biases also 
play a role (\#).   
So far, investigation has explored how worker performance is influenced by
such design factors as the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using alternate problem descriptions
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
the influences that consecutive microtasks exert on one another.

It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend on previous ones.  Such \textit{intertask} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs, with
important implications for the design of crowdsourcing initiatives.

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform, in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
In our first experiment, our prime concepts were \textit{food} and (non-food) 
\textit{objects}.  In our second experiment, the prime concepts were 
\textit{food} and \textit{culture}.  The test tasks for the experiments used 
images that combined both prime concepts used in the initial tasks. Workers 
participated in only one treatment of one experiment, and were randomly 
assigned initial tasks based on one of the 
prime concepts.  From the perspective of the worker, we made no distinction 
between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and experimented with 
requiring the worker to echo the frame using a combo-box input.

We found that intertask effects were stronger than framing, unless the worker
was required to echo the frame, in which case the effects of both exposures were on par.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  Using the wordnet corpus, we analyzed 
the effects on the vocabulary used by workers when 
labeling test tasks. 
Exposure to both framing and initial tasks significantly altered the 
diversity of vocabulary that workers used when referring to the prime concept,
as well as the degree to which workers used specialized (rather than generic) 
words.

Contrary to our initial expectations, our findings suggest that the HPU 
programmer can leverage intertask effects.  By carefully designing workflows,
it is possible to tune worker focus, specificity, and expressivity.  Our 
findings carry two lessons for the task designer: intertask
effects can severely bias HPU output if uncontrolled, but if properly 
controlled, intertask effects might be of significant utility in achieving 
expert-level output.

\paragraph*{Measuring computational hysteresis.}
Before describing the details of our experiments, we explain some 
choices made in measuring intertask and framing effects 
(hereafter collectively \textit{exposure effects}).  
In trying to establish a measure of exposure effects, it might be tempting to 
invoke the notion of a ``neutral frame of mind'' to serve as a baseline. 
But such a concept seems both difficult to define, and unlikely to arise.
Therefore, we always consider the effect of one exposure 
\textit{relative} to another.  We organized our experiments 
into pairs of treatments, between which it is possible to make meaningful 
comparisons.

Our goal in measuring exposure effects is to show that they can actually 
alter the course of algorithms.
We could, by means of a $\chi^2$ test, reject the null hypothesis that
differently exposed HPUs have the same distribution of outputs.
However, this alone would not demonstrate a practically significant effect, in 
terms altering the execution path of algorithms.

We developed a more stringent measure in which we train classifier algorithms 
to infer the exposure of an HPU from its outputs.
If a classifier infers an HPU's exposure with significantly 
better accuracy than provided by chance, then the HPU outputs must have 
been affected by the exposure.  
But in addition, the classifier serves as a kind of certificate, since it is 
an example of an algorithm whose path of execution is influenced by the 
exposure of 
individual HPUs.  In the supplementary material, we ground our approach 
rigorously in terms of the L1-distance metric, which is a measure of 
divergence between distributions.  The L1-distance ranges from 
0 to 1.  In our case, a value of 1 indicates that the HPU's exposure
influences its output so strongly that the exposure can be inferred from the
output without ambiguity.  A value of 0 indicates that the exposure has no
impact on the HPU's outputs.  The connection between the
L1-distance, $D_\mathrm{L1}$, and a classifier's accuracy, $\eta$, is given by:
\begin{equation}
	D_\mathrm{L1} \geq 2 \eta - 1, 
	\label{l1}
\end{equation}
Since a classifier's performance provides a lower bound on the L1-distance, 
we shall denote $D_\mathrm{L1}^-\equiv 2\eta-1$.
The L1-distance admits the following interpretation: if two distributions
$P$ and $Q$ have an L1-distance $d$, then in the worst case, the 
frequency with which a binary algorithm operating on samples
from $P$ outputs a 1, will differ from that when operating on samples from 
$Q$ by $d \times 100\%$.  
Thus, we can interpret $D_\mathrm{L1}^-$ as a kind of
lower bound for the worst-case bias induced by HPU hysteresis.
We provide further discussion of the statistical properties of our approach 
in the supplementary material.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.jpg}
	\caption{
		Examples of images used in labeling tasks. (The italicized treatment 
		names that follow are defined in Table \ref{table:treatments}.) Images 
		from initial tasks: A) \textit{task1:food}, B) 
		\textit{task1:obj}, D) \textit{task2:food}, 
		E) \textit{task2:cult}.  Images from  test tasks: C) \textit{exp1}, 
		F) \textit{exp2}.  The full set of experimental materials is 
		shown in the supplementary material.
	}

	\label{fig:task}
\end{figure}
We performed two experiments, soliciting respectively 2300, and 900 MTurk 
workers to label images depicting food, kitchen- and dining-ware, and cultural
scenes.  In a given experiment, all workers performed the same test
tasks, the output of which was our main dependent variable.  Before performing
the test tasks, workers were subjected to an exposure, via two 
\textit{modalities}: either they completed a set of initial tasks, or 
were exposed to a frame, or both.  As mentioned, from the perspective of the worker, 
there was no distinction or discontinuity between initial and test tasks.

We organized the experiments into \textit{treatment pairs}.  Both treatments in a 
pair involved the same exposure modality, but differed in the
content of the exposure.  In Table 1, we summarize the various experimental
treatments, and introduce a naming convention that we use to refer to 
treatments and treatment pairs.
In the treatment pairs that studied intertask effects (\textit{task1} and
\textit{task2}), workers performed 10 image-labeling tasks, with the first 
five tasks comprising the initial tasks and the last five tasks comprising the
test tasks.

In the first experiment, the content of the exposures was based on the prime
concepts \textit{food} and \textit{objects} (throughout, ``objects'' should be
taken to imply ``non-edible objects'').  So, 
in \textit{task1:food}, the initial tasks contained images depicting food, 
while in \textit{task1:obj}, they contained images depicting objects, 
including kitchen and dining-ware (see Fig~\ref{fig:task}).  Meanwhile, 
for the exposures in \textit{frame1} we indicating that the study was
``funded by the laboratory for the visual perception of \{ Food and Ingredients
$\vert$ Objects and Tools~\}''.  To further strengthen the frame, we 
introduced another pair of framing treatments, \textit{echo}, with stronger
wording: ``The purpose of this study is to understand the visual perception
of \{ Food and Ingredients $\vert$ Objects and Tools~\}'', and required
the worker to echo the purpose of the study by selecting it among other 
options in a combo-box input.

Note that food and objects are both relatively well-defined classes of 
physical entities.  To explore a different conceptual axis, in 
\textit{exp2} we adopted the prime concepts \textit{food} and \textit{culture}.
Compared to food and objects, culture is more abstract, and, rather than 
comprising a class of physical entities, it encapsulates
sets of beliefs, values, and practices.  Examples of images from the initial 
tasks in \textit{task2} are shown in Fig.~\ref{fig:task}, and the full 
collection of 
experimental materials is presented in the supplementary material.  We 
ensured that all workers only participated in one treatment of one experiment.

\setlength{\tabcolsep}{3pt}
\begin{table}[t]
\centering
\begin{tabular}{ c c c c c }
		\hline \noalign{\smallskip}
		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.6cm}{\centering \textbf{Frame}} & \parbox[c]{1.3cm}{\centering \textbf{Initial\\ tasks}}	\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{6}{*}{ \parbox[c]{0.7cm}{ \phantom{XXX} exp1}} 
			& \multirow{2}{*}{task1} & food & none & food\\
			& & obj & none & objects\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame1} & food & food & none\\
			& & obj & objects & none\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{echo} & food & food & none\\
			& & obj & objects & none\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{4}{*}{exp2} 
			& \multirow{2}{*}{task2}  &  food & none & food\\
			& 	&  cult & none & culture\\
			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame2} & food & food & food\\
			& 	& cult & culture & food\\

		\noalign{\smallskip} \hline  
	\end{tabular}

	\caption{ \footnotesize{ 
		We use the treatment names shown to refer to specific sections of 
		the data collected.  For example, \textit{task2:food} refers to the
		treatment in the second experiment which was not exposed to a frame, 
		and whose initial tasks contained images of food (7th row).
	}}
	\label{table:treatments}
\end{table}

\paragraph{The strength of intertask effects.}

We assessed the strength of intertask and framing effects
by training a multinomial naive Bayes classifier to infer a worker's exposure
based on the labels provided in the test tasks (see Fig.~\ref{fig:theta}).
This inference was made over the two possibilities for a given treatment pair.
We found that in \textit{task1}, whether a worker performed food- or 
object-containing initial tasks had a strong effect on their subsequent 
labels, introducing a bias that probably exceeds 30\%.  In comparison, the 
framing effect in \textit{frame1} did 
not influence workers' labels enough for the classifier to make predictions 
with significantly better accuracy than that afforded by chance.  The results 
for \textit{exp2} show a similar, but even more pronounced trend, with the bias
induced by intertask effects in \textit{task2} probably exceeding 50\%.

We see a comparable effect size in the \textit{echo} treatment pair.
We expected \textit{echo} to show a very strong effect because, by requiring
the worker to echo the purported purpose of the task, we signal that it
is of particular importance.  The fact that the intertask effects were on par with the framing 
effects in \textit{echo} shows just how powerful intertask effects can be.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Lower-bound estimates for the  worst-case bias introduced in image
		labeling tasks due to worker exposure effects.  The estimate was 
		based on the performance of a multinomial naive Bayes classifier,
		trained to infer HPU exposures. A) Bias induced through various
		modalities and prime concepts (see Table 1). B) Bias induced in 
		\textit{task1}, as a function of test task position.  Error bars
		show the one-tailed 95\% confidence interval.
	}
	\label{fig:theta}
\end{figure}

\paragraph{Dynamics of intertask effects.} It is natural to expect 
intertask effects to attenuate as a worker proceeds through the test tasks.  
In \textit{task1}, we permuted the test tasks, allowing us to 
measure intertask effects as a function of task position, while factoring
out the content-dependent differences between tasks
(see Fig.~\ref{fig:theta}B). 
We found intertask effects were most pronounced for the first test task,
dropping off substantially for the second, but maintaining a 
significant effect through until the fifth test task.  
This shows that the range of intertask effects extends across at least five 
tasks, and probably substantially more.
This makes intertask effects of greater concern, but as we will show, it also
makes intertask effects of greater potential utility.
In future work, it would be interesting to investigate whether a worker's
susceptibility to intertask effects varies across a large batch.  For example,
It is conceivable that the first few tasks induce especially strong
or long-lasting effects, in effect ``setting the tone'' for the rest of the batch.

\paragraph{Detailed influences on worker outputs.} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/vocab_specificity.pdf}
	\caption{
		Detailed exposure effects on workers' label vocabulary, for various
		exposure modalities and prime concepts (see Table 1).  A) Change in 
		the number of food-related labels used. B) Change in the total number 
		of \textit{unique} food-related labels used. C) Relative specificity
		of treatment pairs, when considering food-related labels.  For all
		plots, a positive bar height indicates an increase for the 
		food-exposed treatment for the pair indicated on the abscissa.
	}
	\label{fig:specificity}
\end{figure}

Having observed significant effects from framing and task-exposure,
we looked in greater detail at the nature of the effects.  For instance,
does a food-oriented exposure induce workers to provide more food-related
labels?  And if so, does this explain the classifier's ability to distinguish
a worker's initial exposure?

To answer these questions, we used the wordnet corpus, augmented with names of
ethnic foods learned from the Internet.  Wordnet provides a (roughly)
hierarchical set of relationships between nouns, called hypernym-hyponym
relationships.  A hypernym is a generalization, while a hyponym is a 
specialization.  So, for example, ``bread'' is hypernym to ``rye bread'' 
and hyponym to ``baked goods''.

Using the augmented hypernym-hyponym structure of wordnet to identify nouns 
depicting food, we found that, for the experimental pairings in 
\textit{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain the classifier's ability to 
infer worker exposures, it is insufficient to explain the observed performance
(*).
Furthermore, in \textit{exp2:task2} we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A), which suggests a more complicated 
relationship between exposure and label usage.
Given the difference between the food-culture and food-object conceptual axes, 
it should perhaps not be surprising that the effect of initial task exposure 
is different.  It would appear that we must be prepared to accommodate 
multiple 
countervailing factors influencing the number of references to the prime 
concept (food).  We will come back to this point momentarily.
